### 2016-12-06 introduction.tex
Restructured chapter1.tex as introduction.tex

% Chapter 1: Introduction

@@

The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.  Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.  The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.  The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.  This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bachivan2010}.  This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.  AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.  The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.  In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.

@@

There have been at least 6 mechanisms proposed for how polar patches form on the dayside \citep{Crowley1996,Carlson2012}.
\begin{enumerate}
	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
	\item Average flow patterns varying in time \citep{Anderson1988}
	\item Transient magnetopause reconnection \citep{Lockwood1992b}
	\item Plasma production by cusp particle precipitation \citep{Rodger1994,Millward1999}
	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
	\item Alfven wave coupling \citep{Prikryl1999}
\end{enumerate}
Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.  These high densities must originate from the resivor of solar illuminated plasma on the dayside.  Additionally, a mechanism is still required to separate patches from the cusp region.  High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.  Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.  However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.

@@

By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.  First, assume a cold plasma such that the temperature of both ions and electrons is negligible.  In this case, \(T_\alpha = 0\), so \(C_\alpha = 0\).  Secondly assume all irregularities are perfectly aligned with the magnetic field such that \(k_\parallel = 0\).  This means that \(\vec{k}_\perp = \vec{k}\), \(y=0\), and \(Y=0\).  Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \gg \nu_\alpha\)).  If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
\begin{equation}
	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
\end{equation}
Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
\begin{equation}
	\label{eqn:gdi_Mak16_Freg}
	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
\end{equation}
The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
\begin{equation}
	\label{eqn:gdi_Mak16_gam}
	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
\end{equation}
The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.  Using the definitions of \(Z_1\) and \(Z_2\) provided in Equation \ref{eqn:gdi_Mak16_Z} and the simplifying assumptions discussed above for the F region,
\begin{equation}
	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
\end{equation}.
Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.  Additionally the local approximation \(G \ll k_\perp\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
\begin{equation}
	\frac{Z_2}{Z_1} = R+ib(1+R^2)
\end{equation}
Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
\begin{equation}
	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
\end{equation}
Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevitch2014c}.
\begin{equation}
	\label{eqn:gdi}
	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
\end{equation}

@@

This asymmetry was first observed by \citet{Weber1984}.  Backscatter power from HF radars tends to be greater on the trailing edges of both moving high-density polar structures \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.  Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.  Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004}.  \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.  This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients (that observed by ASIP and ASI methods) and structuring was greatest where GDI was operational.

@@

>>>>

% Chapter 1: Introduction
\label{sec:introduction}

@@

The first ISR was built in Arecibo, Puerto Rico in the 1950s \citep{Gordon1958}.  Since then, at least 8 other ISRs have been deployed around the world at equitoral, mid-, and high-latitudes.  The most recent advancement has been the development of Advanced Modular Incoherent Scatter Radars (AMISR), of which there are currently three in use, one at the Poker Flat Rocket Range, just north of Fairbanks, AK and two in Resolute Bay, Canada.  The advantage of these new systems is that they are electronical steerable, unlike older ISRs which consisted of a large dish that had to be moved manually to change the beam direction.  This allows beams to be transmitted in many different directions at a very high time cadence, giving the radar the capability to make measurements in multiple directions almost simultaneously \citep{Nicolls2007a,Nicolls2007b,Bahcivan2010}.  This is very important for creating 2D maps of ionospheric conditions and tracking how they change in time \citep{Semeter2009,Dahlgren2012a,Dahlgren2012b}.  AMISR systems typically have two types of pulses, a long pulse (LP) with 72 km range gates for F-region studies and an alternating code pulse (AC) with 4.5 km range gates gates for E region studies.  The exact number and configuration of beams depends on the radar mode used and can be changed easily due to the system's electronic steering.  In the common WorldDay mode, there are typically 11 beams and data is collected at \(\sim\)1 min intervals, although this can change slightly between different renditions of the WorldDay mode.

@@

There have been at least 6 mechanisms proposed for how polar patches form on the dayside \citep{Crowley1996,Carlson2012}.
\begin{enumerate}
	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
	\item Average flow patterns varying in time \citep{Anderson1988}
	\item Transient magnetopause reconnection \citep{Lockwood1992b}
	\item Plasma production by cusp particle precipitation \citep{Rodger1994,Millward1999}
	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
	\item Alfven wave coupling \citep{Prikryl1999}
\end{enumerate}
Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Rodger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.  These high densities must originate from the resivor of solar illuminated plasma on the dayside.  Additionally, a mechanism is still required to separate patches from the cusp region.  High speed flow channels can ``cut'' segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.  Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.  However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.

@@

By considering certain limiting cases, an expression that is more relevant to the polar F region specifically can be resolved from Equation \ref{eqn:gdi_Mak16_alt}.  First, assume a cold plasma such that the temperature of both ions and electrons is negligible.  In this case, \(T_\alpha = 0\), so \(C_\alpha = 0\).  Secondly assume all irregularities are perfectly aligned with the magnetic field such that \(k_\parallel = 0\).  This means that \(\vec{k}_\perp = \vec{k}\), \(y=0\), and \(Y=0\).  Finally, consider only low frequencies (\(\omega \ll \Omega_\alpha\)) and long wavelengths (\(\vec{k}\cdot\vec{V}_{\alpha 0} \gg \nu_\alpha\)).  If \(D_\alpha\) is expressed as shown below, it can bee seen that these limits result in the first two terms being negligible such that \(D_\alpha \approx r_\alpha\).
\begin{equation}
	D_\alpha = -i\frac{\omega}{\Omega_\alpha} + i\frac{\vec{k}\cdot\vec{V}_{\alpha 0}}{\Omega_\alpha} + \frac{\nu_\alpha}{\Omega_\alpha} \approx \frac{\nu_\alpha}{\Omega_\alpha} = r_\alpha
\end{equation}
Equation \ref{eqn:gdi_Mak16_alt} can then be expressed as follows.
\begin{equation}
	\label{eqn:gdi_Mak16_Freg}
	(r_i-r_e)\omega = \vec{V}_d\cdot\vec{k}\left(\frac{Z_2}{Z_1}\right)+(r_i\vec{V}_{i0}-r_e\vec{V}_{e0})
\end{equation}
The growth rate, \(\gamma\), is contained within the dispersion relation as \(\omega = \omega_r+i\gamma\), so the imaginary part of Equation \ref{eqn:gdi_Mak16_Freg} must be considered separately.
\begin{equation}
	\label{eqn:gdi_Mak16_gam}
	(r_i-r_e)\gamma = \vec{V}_d\cdot\vec{k}\;\text{Im}\left\{\frac{Z_2}{Z_1}\right\}
\end{equation}
The differential plasma drift has been defined previously in Equation \ref{eqn:gdi_Mak16_Vd} but the imaginary component of \(Z_2/Z_1\) must still be identified.  Using the definitions of \(Z_1\) and \(Z_2\) provided in Equation \ref{eqn:gdi_Mak16_Z} and the simplifying assumptions discussed above for the F region,
\begin{equation}
	\frac{Z_2}{Z_1} = \frac{R(i+a)-b}{i+a+Rb}
\end{equation}.
Multiplying both the numerator and denominator by \(-i+a+Rb\) removes the imaginary component from the denominator.  Additionally the local approximation \(G \ll k_\perp\) is taken, which results in \(a\), \(b\), and \(c\) being small so that any second order terms in these quantities can be neglected.
\begin{equation}
	\frac{Z_2}{Z_1} = R+ib(1+R^2)
\end{equation}
Using the above result and Equation \ref{eqn:gdi_Mak16_Vd}, Equation \ref{eqn:gdi_Mak16_gam} becomes
\begin{equation}
	(r_i-r_e)\gamma = (r_i-r_e)s_es_i\left(1+\psi\right)\left(R\vec{V}_E-\frac{E}{B}\right)\cdot\vec{k}b\left(1+R^2\right)
\end{equation}
Recognizing that \(s_es_i\left(1+R^2\right) = (1+\psi)^{-2}\) and \(b = \vec{G}\cdot\vec{k}\times\uvec{b}/k^2 = -\vec{k}\cdot\uvec{b}\times\vec{G}/k^2\) by vector identities, a final expression for the GDI growth rate in the F region can be found, which agrees well with the results of \citet{Makarevich2014c}.
\begin{equation}
	\label{eqn:gdi}
	\gamma = \frac{1}{1+\psi}\left(\uvec{k}\cdot\uvec{b}\times\vec{G}\right)\uvec{k}\cdot\left(\frac{\vec{E}}{B}-R\vec{V}_E\right)
\end{equation}

@@

This asymmetry was first observed by \citet{Weber1984}.  Backscatter power from HF radars tends to be greater on the trailing edges of both moving high-density polar structures \citep{Milan2002b} and sun-aligned arcs identified from ASIP data \citep{Koustov2012}, which is generally attributed to GDI being unstable on the trailing edge but stable on the leading edge of such large, moving density structures.  Studies of patches using all-sky airglow imagers (ASI) found that the density gradient on the leading edge of polar patches tends to be 2--3 times steeper than that on the trailing edge \citep{Hosokawa2016}.  Additionally, large finger-like structures on the trailing edge of patches have been identified that were 10s--100s of kilometers in size and agree with predictions made by GDI simulations \citep{Gondarenko2004b}.  \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.  This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients (that observed by ASIP and ASI methods) and structuring was greatest where GDI was operational.

@@

\bibliographystyle{plainnat}
\bibliography{references}
